Constraints on magic state protocols from the statistical mechanics of Wigner negativity

### 2022-01-04 npj_revisions/referee_responses.tex

\documentclass[11pt]{letter}

\usepackage[colorlinks=true,urlcolor=blue, hyperindex,breaklinks=true] {hyperref}

\usepackage{graphicx}
\usepackage[a4paper,left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{fancyhdr}
\usepackage[table,xcdraw,dvipsnames]{xcolor}

@@

\def\A{{\cal A}}
\def\Z{{\cal Z}}
\def\H{{\cal H}}
\def\E{{\cal E}}
\def\J{{\cal J}}
\def\R{{\cal R}}
\def\D{{\cal D}}
\def\M{{\cal M}}
\def\F{{\cal F}}
\renewcommand{\O}{{\cal O}}
\renewcommand{\P}{{\cal P}}

\begin{document}

Dear Prof. Sven Rogge,

Thank you for passing on the Reviewers' evaluation of our work. We are very glad that both Reviewers had positive opinions on it, and raised thought-provoking questions for future work.

Below we address the comments of both Reviewers, and respond to questions raised.  The revisions we have made are highlighted in the re-submitted version of the manuscript.\\

@@

\textit{They also discuss the case of finite temperature, which shows that their method can be generalized to a variety of situations.}

\textit{They also show that one can define R\'{e}nyi entropies for quasi-probability distributions, and derive an elegant interpretation of the mana as the residue of a pole quantifying the divergence of the R\'{e}nyi entropy as it approaches the Shannon entropy.}

\textit{Overall, the results of the paper appear to me to be novel, topical, and mathematically elegant. I also found the paper to be well-written. I therefore strongly recommend it for publication.}

We thank the Reviewer for the positive comments on our work.

@@

\textit{However, I have one suggestion that could be included in a minor revision:}

\textit{The main result of the authors seems to rest on Theorem 1, which in turn relies on the unnumbered result on page 5, after equation (15), which states roughly that “relative majorization holds between pairs of probability distributions holds if the Lorenz curve of one is greater than or equal to the Lorenz curve of the other, for all x.” The authors refer to [103] for a proof of this fact. However, I think it would substantially enhance accessibility and readability if the authors could give a proof of this result within the present paper, perhaps in an appendix.}

We are very grateful for suggestions on making our results more accessible -- we have added a proof of the result in Supplementary Note~2 and a reference to it in the main text.

@@

\textit{A side remark: As I understand, Lorenz curves were initially defined in economics to describe the income distribution of a population. It seems possible for some individuals to lose money during the year, so the distribution of incomes in a population is in general, more analogous to a quasi-probability distribution rather than a probability distribution. So, perhaps some of the basic results of the paper may also be applicable, in principle, to a wider context in economics.}

This is an interesting idea,  and following the Reviewer's suggestion we have thought about it a little.  It seems that variants of the Gini coefficient do exist to account for losses,  however the level of technical detail is quite simple.  The R\'{e}nyi entropies we consider here would certainly give finer information about levels of inequality,  although the physical or statistical interpretations are not so clear to us at present.  We might expect that the $\alpha \rightarrow 1$ entropy measures the total cumulative losses within a country,  in the same way as mana for magic states relates to the sum negativity.

In traditional majorization this kind of thinking has occurred previously -- for example Robin-Hood transforms that take from the rich and give to the poor. Of course in the vein of the present paper,  one could construct a ``resource theory of social inequality'' (although calling this a ``resource'' is questionable,  except perhaps for billionaire CEOs) in which socialist policies that redistribute wealth more fairly over a population are the ``free/easy operations'' that transform a country.  Distillation protocols on multiple countries would then address the question of the degree to which one could form high levels of inequality in one country solely via collective socialist actions on multiple countries.  However, we do not pursue such social engineering theories any further here.

@@

\textit{Interestingly, previous metrics, such as mana proposed by Veitch et al., can be obtained from this more fundamental quantity. Authors then proceed to show that this realization can be used to upper bound the rate of magic transformation in the context of strange state, a particular magic state for qutrits. They also study limits of magic transformation by providing each qudit with a Hamiltonian and understanding temperature dependent transformations.}

\textit{This is a strong result. However, before accepting for publication, I would like authors to address some comments.
Questions/comments for the authors:}

\textit{1. In section 3C, an upper bound on the rate of distillation for qutrit magic states is presented. However, we do not see an explicit scheme that can actually achieve this bound. Some comments on whether this bound is achievable would be appreciated.}

All existing distillation protocols have rates that are very far from any general upper bounds that have been derived e.g.  via resource monotones. While ours improve on prior upper bounds, there still remains a gap between known rates and our bounds. Therefore, we do not know if our upper bounds can be achieved, except in trivial scenarios. Since our work captures the constraints due to stochasticity in the Wigner representation, the gap between realisable rates and our bounds implies that the (symplectic) phase space structure is crucial in determining what actual rates can be obtained.

@@

\textit{In the conclusions, short discussion on lower bounds can be lengthened. Does this require explicit error correcting codes for distillation?}

The proposed route to lower bounds would be separate from the method of explicit distillation rates from error correcting codes.  Instead, the proposal is to leverage the critical additional structure that the Wigner distribution carries,  above and beyond involving stochasticity.  The upper bounds obtained are upper bounds on distillation rates because the set of stochastic maps is a superset for the actual maps that free operations generate.  
Therefore, one can approach rates from the other direction: construct a set of stochastic maps that respect the structure of a subgroup of maps that correspond to free operations and then compute the associated majorization constraints for this restricted set. This would provide lower bounds on what free operations can achieve.

@@

The difficulty in obtaining lower bounds lies in the choice of sensible sets of maps. 
We have briefly explored the stochastic maps one gets from Weyl covariant channels -- these can be shown to be simply represented as convex mixtures of unitary displacements and in fact one can solve the majorization conditions for this class of channels (it is called cyclic majorization in the mathematical literature~[136]). The problem is that while Weyl covariant channels are a nice family of channels,  they turn out to be useless for distillation (which can be shown via majorization). 
More non-trivial and useful classes of operations could be analysed but the problem becomes that the majorization conditions may not be solvable in closed form (we flag a class of operations based on finite-reflection groups~[134] -- they are known to have finite conditions for majorization with respect to a discrete group). Beyond tidy closed analytics, one can use techniques from semi-definite programming to estimate distillation rates -- which might be an interesting future project to explore.

We have extended our discussion in the main text with some of the above comments.

\textit{2. It was shown by Bengtsson et al. that qudit magic states for prime-dimensional qudits, where the prime p is 1 mod 3, there exist distinct magic states under distinct orbits of the Clifford group. In other words, there exist magic states such that we cannot convert from one to other using only Clifford operations. Can authors reproduce this result using their theory?}

This is another interesting suggestion.  We have done initial calculations on this,  and in principle we could reproduce their results, however to do so would require a more detailed analysis. Specifically,  we have computed the Lorenz curves for the pairs of states relative to different probability distributions $\mathbf{r}$ on the phase space. In all situations considered the Lorenz curves intersect each other which implies the impossibility of interconversion under channels leaving the distribution $\mathbf{r}$ invariant.  To complete the proof one would have to prove that this occurs for \emph{all} choices of $\mathbf{r}$. 
%In the case of stabilizer \& Clifford circuits, one can consider $\mathbf{r}$ obtained from stabilizer states and in fact pure stabilizer states would suffice. 
%If one varies over all $\mathbf{r}$ corresponding to the free states of the theory, then the completeness result given in Theorem 3 of the revised manuscript implies the impossibility of the interconversion. 
Varying over all positive $\mathbf{r}$ would in fact imply a stronger result: the impossibility of interconversion under quantum channels with positive Wigner representations.

@@

\textit{3. Perhaps authors have already mentioned this, but Clifford operations can also be supplemented by Pauli measurements and adaptive gates. In the main text, as well as in conclusions (comments about G-majorization) this is not mentioned. Would this lead to distinct/ refined bounds?}

This is a good point -- these operations still admit representations in terms of positive, stochastic maps and therefore our results cover this regime, and the bounds remain unchanged. 
For example, according to Jozsa and Van den Nest's work [Jozsa R and Van den Nest M (2013) Classical simulation complexity of extended Clifford circuits, arXiv:1305.6190], intermediate Pauli measurements and adaptive Clifford gates on $N$ qubits can always be substituted by a Clifford gate on $(N+k)$ qudits, with $k \in \mathbb{N}$.
Therefore, our bounds are directly applicable on the equivalent circuit.

We do agree that it is important to better engage the actual classes of operations of interest from a quantum computing perspective, and in recent work we are extending the approach to qubit systems. There, it is much more important to be clear on what actual operations are permitted and we engage deeper with such concrete techniques in quantum computing (in particular CSS-code based protocols).

@@

The dimension of the system does arise in the bounds in fairly simple ways (for example see the R\'{e}nyi entropy bounds for unital operations in Eq.(52) of the main text).  However it does not depend on prime properties of the dimension.  The only non-trivial dependence on primality is in the original Wigner representation,  with the $p=2$ prime case famously having a more complex structure.  We might expect that if one consider $G$-majorization for the lower bounds,  one obtains non-trivial features that do depend on the particular prime dimension.  In this vein,  we note that the symplectic groups for $p=3$ and $p=5$ are finite reflection groups and therefore the associated $G$-majorization for each case is fully solvable in terms of a finite number of conditions.

\textit{5. In optics, the Wigner function is not the only quasi-probability distribution. Indeed, there are the P and Q distributions and potentially a continuous spectrum of quasiprobability distributions. In the discrete setting the Wigner function occupies a special role because of Gross' work showing that stabilizer states can be represented positively for prime-power dimensions and other useful properties like Clifford covariance. However is it possible that the other distributions offer extra information not captured by the Wigner distribution? Can we use a discrete analogue of the P and Q distributions?}

We again thank the Reviewer for this nice point. In our paper, we first establish results for \emph{arbitrary} quasi-distributions,  and only then specialise to Wigner quasi-distributions. The Supplementary Material has general analysis for arbitrary quasi-distributions, which hopefully makes this clear.

The question of whether switching representations provides extra information is a very interesting one,  and it is not obvious to us when this occurs.  In the context of magic distillation the central thing that must be checked is that the free operations are postively represented as stochastic maps in this new representation -- if they are not,  then the majorization theory cannot be applied.  However, if the free operations are still positively represented then we can certainly imagine situations in which the majorization bounds can be improved.  Perhaps a good strategy would be to look for a representation that has the largest contrast between free operations and the target magic state: i.e.  all free operations are positively represented,  but the magic state has very ``large'' negativity in this representation.  This seems to us an interesting line to explore,  perhaps using the discrete analogues to the $s$--ordered Cahill/Glauber distributions due to Ruzzi \textit{et al.} [Ruzzi M \textit{et al.} (2005),  J. Phys. A: Math. Gen. 38 6239].  Or more generally from frame-theory, by perturbing the Wigner frame defined via $\{D_{\x}A_0 D_{\x}^\dagger\}$ to the more general scenario of $\{D_{\x} M D_{\x}^\dagger\}$ where $M$ is an arbitrary trace-one operator that can be tuned to vary the choice of frame,  while still keeping covariance properties.

### 2022-01-04 magic_constraints_maintext_numbered.tex
revised manuscript

\newcommand{\revhigh}[1]{{\color{red}#1}}
\hypersetup{
	colorlinks=true,  
	linkcolor=blue,   
	citecolor=blue,   
	urlcolor=blue     
}

@@

\begin{abstract}
Magic states are key ingredients in schemes to realize universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization to describe Wigner negative magic states for qudits of odd prime dimension processed under Clifford circuits. We show that majorization allows us to both quantify disorder in the Wigner representation and derive upper bounds for magic distillation. These bounds are shown to be tighter than other bounds, such as from mana and thauma, and can be used to incorporate hardware physics, such as temperature dependence and system Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We find that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting. \revhigh{REMOVE (NO CLAIMS TO NOVELTY): This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises novel questions in the context of classical statistical mechanics.}
\end{abstract}

@@

\subsection*{Overview of our results}
\revhigh{
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations. We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels. These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed-point structure. We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that enable a simple description in the thermodynamic limit before giving entropic analysis that can be applied generally.}

The relations between the sets of protocols considered are as follows:
\begin{align}
\mbox{Unital} &\subset \mbox{ local fixed-points} \nonumber\\ 
&\subset \mbox{ sub-linear correlations} \subset \mbox{ general protocols.} \nonumber
\end{align}

@@

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.38]{distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorization on unital protocols. The constraints are plotted as a function of the depolarizing error $\epsilon$ for a noisy Strange state (Eq.~(\ref{eq:noisysn})). \revhigh{The bound $R_\infty$ follows from a Lorenz curve analysis presented before Eq.~(\ref{eq:unital-bound}), while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in the derivation of Eq.~(\ref{eq:Ra})}. Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana~\cite{cit:veitch2} and max--thauma $\theta_{\rm{max}}$~\cite{Wang_2020} bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\revhigh{
\subsection*{Phase space representation of quantum mechanics}
}

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\P_d \coloneqq \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\z \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\z} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend the representation to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\H_{d_1} \otimes \dots \otimes \H_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\z} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\z &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \P_{d_1} \times \dots \times \P_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\z \in \mathbb{Z}_d^{2n}$.

@@

In the usual case where $\w$ and $\r$ are both probability distributions the Lorenz curve is defined on the interval $[0,1]$, and rises monotonically until it reaches the value $1$ at $x=1$. The value $L_{\w|\r}(x) = 1$ is a global maximum. Moreover, if $\w, \w', \r, \r'$ are all valid probability distributions with $\r,\r'$ having positive components, then
\begin{equation*}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } L_{\w |\r}(x) \ge L_{\w' |\r'}(x),
\end{equation*}
for all $x \in [0,1]$. 
\revhigh{This result is shown in~\cite{ruch_mixing_1978} and we reproduce it in Supplementary Note~2.}
It provides a simple way of computing whether relative majorization holds between pairs of probability distributions.

@@

\begin{theorem}\label{thm:lcquasi}
	Let $\w \in \mathbb{R}^n,\w'\in \mathbb{R}^m$ be quasi-distributions, and let $\r \in \mathbb{R}^n, \r' \in \mathbb{R}^m$ be probability distributions with non-zero components. Then, $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w|\r}(x) \ge L_{\w' |\r'}(x)$ for all $x \in [0,1]$.
\end{theorem}
\begin{proof}
	Since the components of $\r$ are strictly positive, there always exists an $\epsilon >0$ such that $\w_\epsilon \coloneqq \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. A similar result holds for $\w'$ and $\r'$ and we choose $\epsilon$ sufficiently small so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. We now have that $(\w, \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if 
\begin{equation}
A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'\mbox{ and } A\r = \r'.
\end{equation}
In terms of a Lorenz curve condition we have that $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$ if and only if $L_{\w_\epsilon |\r} (x) \ge L_{\w'_\epsilon |\r'} (x)$ for all $x \in [0,1]$. 
Additionally, we show in Supplementary Note~2 that the Lorenz curve for any quasi-distribution obeys the relation
\begin{equation}
L_{a \w + b \r | \r} (x) = a L_{\w|\r} (x) + bx,
\end{equation}
for any $a >0$ and $b \in \mathbb{R}$. This relation implies that
\begin{align*}
L_{\w_\epsilon |\r} (x) &\ge L_{\w'_\epsilon |\r'} (x) \nonumber \\ 
&\mbox{ if and only if }& \\
\epsilon L_{\w |\r} (x) + (1-\epsilon) x &\ge \epsilon L_{\w' |\r'} (x) + (1-\epsilon) x.
\end{align*}
Finally, the $(1-\epsilon)x$ terms cancel on both sides and we get the required relative majorization conditions that $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w | \r} (x) \ge L_{\w' | \r'}(x)$ for all $x \in [0,1]$, as required.
\end{proof}

@@

We can now consider how majorization constrains magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure, which allows the use of $\d$--majorization. 
\revhigh{Later, we drop the requirement of fixed-points and instead use the full relative majorization, before giving general entropic constraints.}

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, we define the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those free operations that leave the distinguished state $\sigma \in \F$ invariant.
\end{definition}
This gives a simple way to break up any theory into smaller, more manageable parts, and where the union over all sub-theories returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho_1 \rightarrow \rho_2$ is possible in $\R$ if and only if the transformation $\rho_1 \rightarrow \rho_2$ is possible in at least one sub-theory $\R_\sigma$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\R_\sigma$ via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer fixed-point theorem~\cite{cit:brouwer}, this mapping must therefore have a fixed point $\sigma \in \F$, so $\E \in \O_\sigma$ and the interconversion is possible in $\R_\sigma$.
\end{proof}
We can compare with resource monotones, which are also called resource measures. A complete set of measures $\{\M_i\}_i$ is such that $\rho \rightarrow \sigma$ if and only if $\M_i(\rho) \ge \M_i(\sigma)$ for all $i$. The above set of sub-theories can therefore be viewed as a complete set of ``co-measures'' for the theory, where the parent resource theory's complex pre-order of states is mapped to a simpler pre-order for a particular $\R_\sigma$. 
In the next section we show that a given $\R_\sigma$ can be approximated by a majorization pre-order, and from this we compute magic distillation bounds.

@@

We consider some $\sigma \in \F$, and its corresponding $\R_\sigma$. We are interested in the ability to transform many copies of some noisy magic state $\rho$ towards a more pure form of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\z) < 0$ for some regions of $\z \in \P_d$. The state $\sigma$ is assumed to have a Wigner distribution with $W_\sigma(\z) > 0$ for all $\z$ in the phase space.
\revhigh{This is justified because a non-full rank state $\sigma$ can be handled as a limiting case in which we first add an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$ and then take $\epsilon \rightarrow 0$.}

@@

Note that this is not a single numerical constraint, but a family of constraints. For $n$ copies of a qudit system of dimension $d$ the number of terms in $W_{\rho}$ is $d^{2n}$, so imposing the Lorenz curve condition corresponds to exponentially many constraints. 
%Before computing explicit examples, in the next section we describe how these majorization constraints relate to other known monotones.

%\subsection*{Basic monotones for magic distillation for a particular $\R_\sigma$ and the relation to mana}

>>>>

\hypersetup{
	colorlinks=true,  
	linkcolor=blue,   
	citecolor=blue,   
	urlcolor=blue     
}

\usepackage{lineno}
\linenumbers

@@

\begin{abstract}
Magic states are key ingredients in schemes to realize universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization to describe Wigner negative magic states for qudits of odd prime dimension processed under Clifford circuits. We show that majorization allows us to both quantify disorder in the Wigner representation and derive upper bounds for magic distillation. These bounds are shown to be tighter than other bounds, such as from mana and thauma, and can be used to incorporate hardware physics, such as temperature dependence and system Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We find that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting.  This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises non-trivial questions in the context of classical statistical mechanics.
\end{abstract}

@@

\subsection*{Overview of our results}
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations. We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels. These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed-point structure. We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that enable a simple description in the thermodynamic limit before giving entropic analysis that can be applied generally.

The relations between the sets of protocols considered are as follows:
\begin{linenomath}\begin{align}
	\mbox{Unital} &\subset \mbox{ local fixed-points} \nonumber\\ 
&\subset \mbox{ sub-linear correlations} \subset \mbox{ general protocols.} \nonumber
\end{align}\end{linenomath}

@@

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.38]{distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorization on unital protocols. The constraints are plotted as a function of the depolarizing error $\epsilon$ for a noisy Strange state (Eq.~(\ref{eq:noisysn})). The bound $R_\infty$ follows from a Lorenz curve analysis presented before Eq.~(\ref{eq:unital-bound}), while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in the derivation of Eq.~(\ref{eq:Ra}). Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana~\cite{cit:veitch2} and max--thauma $\theta_{\rm{max}}$~\cite{Wang_2020} bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\subsection*{Phase space representation of quantum mechanics}

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{linenomath}\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}\end{linenomath}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\P_d \coloneqq \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\z \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\z} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend the representation to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\H_{d_1} \otimes \dots \otimes \H_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\z} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{linenomath}\begin{align*}
	\z &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \P_{d_1} \times \dots \times \P_{d_n},
\end{align*}\end{linenomath}
to denote the phase space point for the composite system. 
For simplicity, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\z \in \mathbb{Z}_d^{2n}$.

@@

In the usual case where $\w$ and $\r$ are both probability distributions the Lorenz curve is defined on the interval $[0,1]$, and rises monotonically until it reaches the value $1$ at $x=1$. The value $L_{\w|\r}(x) = 1$ is a global maximum. Moreover, if $\w, \w', \r, \r'$ are all valid probability distributions with $\r,\r'$ having positive components, then
\begin{linenomath}\begin{equation*}
	(\w, \r) \succ (\w', \r') \mbox{ if and only if } L_{\w |\r}(x) \ge L_{\w' |\r'}(x),
\end{equation*}\end{linenomath}
for all $x \in [0,1]$. 
This result is shown in~\cite{ruch_mixing_1978} and we reproduce it in Supplementary Note~2.
It provides a simple way of computing whether relative majorization holds between pairs of probability distributions.

@@

\begin{theorem}\label{thm:lcquasi}
	Let $\w \in \mathbb{R}^n,\w'\in \mathbb{R}^m$ be quasi-distributions, and let $\r \in \mathbb{R}^n, \r' \in \mathbb{R}^m$ be probability distributions with non-zero components. Then, $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w|\r}(x) \ge L_{\w' |\r'}(x)$ for all $x \in [0,1]$.
\end{theorem}
\begin{proof}
	Since the components of $\r$ are strictly positive, there always exists an $\epsilon >0$ such that $\w_\epsilon \coloneqq \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. A similar result holds for $\w'$ and $\r'$ and we choose $\epsilon$ sufficiently small so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. We now have that $(\w, \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if 
\begin{equation}
A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'\mbox{ and } A\r = \r'.
\end{equation}
In terms of a Lorenz curve condition we have that $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$ if and only if $L_{\w_\epsilon |\r} (x) \ge L_{\w'_\epsilon |\r'} (x)$ for all $x \in [0,1]$. 
Additionally, we show in Supplementary Note~2 that the Lorenz curve for any quasi-distribution obeys the relation
\begin{equation}
L_{a \w + b \r | \r} (x) = a L_{\w|\r} (x) + bx,
\end{equation}
for any $a >0$ and $b \in \mathbb{R}$. 
This relation implies that $L_{\w_\epsilon |\r} (x) \ge L_{\w'_\epsilon |\r'} (x)$ if and only if $\epsilon L_{\w |\r} (x) + (1-\epsilon) x \ge \epsilon L_{\w' |\r'} (x) + (1-\epsilon) x$. 
Finally, the $(1-\epsilon)x$ terms cancel on both sides and we get the required relative majorization conditions that $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w | \r} (x) \ge L_{\w' | \r'}(x)$ for all $x \in [0,1]$, as required.
\end{proof}

@@

We can now consider how majorization constrains magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure, which allows the use of $\d$--majorization. 
Later, we drop the requirement of fixed-points and instead use the full relative majorization, before giving general entropic constraints.

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, we define the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those free operations that leave the distinguished state $\sigma \in \F$ invariant.
\end{definition}
This gives a simple way to break up any theory into smaller, more manageable parts, and where the union over all sub-theories returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho_1 \rightarrow \rho_2$ is possible in $\R$ if and only if the transformation $\rho_1 \rightarrow \rho_2$ is possible in at least one sub-theory $\R_\sigma$.
\end{theorem}
\noindent The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\R_\sigma$ via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer fixed-point theorem~\cite{cit:brouwer}, this mapping must therefore have a fixed point $\sigma \in \F$, so $\E \in \O_\sigma$ and the interconversion is possible in $\R_\sigma$.
\end{proof}
We can compare with resource monotones, which are also called resource measures. A complete set of measures $\{\M_i\}_i$ is such that $\rho \rightarrow \sigma$ if and only if $\M_i(\rho) \ge \M_i(\sigma)$ for all $i$. The above set of sub-theories can therefore be viewed as a complete set of ``co-measures'' for the theory, where the parent resource theory's complex pre-order of states is mapped to a simpler pre-order for a particular $\R_\sigma$. 
In the next section we show that a given $\R_\sigma$ can be approximated by a majorization pre-order, and from this we compute magic distillation bounds.

@@

We consider some $\sigma \in \F$, and its corresponding $\R_\sigma$. We are interested in the ability to transform many copies of some noisy magic state $\rho$ towards a more pure form of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\z) < 0$ for some regions of $\z \in \P_d$. The state $\sigma$ is assumed to have a Wigner distribution with $W_\sigma(\z) > 0$ for all $\z$ in the phase space.
This is justified because a non-full rank state $\sigma$ can be handled as a limiting case in which we first add an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$ and then take $\epsilon \rightarrow 0$.

@@

Note that this is not a single numerical constraint, but a family of constraints. For $n$ copies of a qudit system of dimension $d$ the number of terms in $W_{\rho}$ is $d^{2n}$, so imposing the Lorenz curve condition corresponds to exponentially many constraints.

Before computing explicit examples, we state some generic aspects of Lorenz curves for magic states, which allow us to interpret previous magic monotones in a new light. The first result gives a simple way to see that the sum-negativity/mana of a magic state is a monotone~\cite{cit:veitch2}.
